Modular Model-Checking Implementations of a Byzantine Fault-Tolerant Protocol

### 2016-12-07 doc/om-paper/paper.tex
more edits!

\usepackage{ifthen}
\usepackage{xcolor}

\usepackage{graphicx}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{multirow}
\usepackage{multicol}

@@

\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}}
\newcommand{\OMH}{\ensuremath{\mathrm{OMH}}}
\newcommand{\Id}{\ensuremath{\mathrm{Id}}}
\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}}
\newcommand{\ERR}{\ensuremath{\mathrm{ERR}}}
\newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}}

@@

With this machinery, we can define the semantics of faults by constraining the relationship between a message broadcast and the values received by the recipients. For a nonfaulty process that broadcasts, every recipient receives the sent message, and for symmetric faults, there is no requirement that the messages sent are the ones received, only that every recipient receives the same value:
\noindent
\begin{multicols}{2}
\begin{align*}
&nonfaulty\_constraint =\\
  &\quad \forall i, j \in \Id, t \in \mathbb{R}.\\
  &\quad\quad faults(i) = none\\
  &\quad\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
\end{align*}\break
\noindent
\begin{align*}
&sym\_constraint =\\
  &\quad \forall i, j, k \in \Id, t \in \mathbb{R}.\\
  &\quad\quad\quad (\phantom{\textnormal{and }} faults(i) = sym\\
  &\quad\quad\quad \phantom{(}\textnormal{and } broadcast(i, \{j, k\}, t, m))\\
  &\quad\quad \textnormal{ implies } recv(i, j, t) = recv(i, k, t)
\end{align*}
\end{multicols}
\noindent
Byzantine faults are left completely unconstrained.

@@

As a point of comparison, we also benchmarked Rushby's model~\cite{Rushby:OM1} on the same machine, using the same environment. When the number of receivers was fixed at 2, the verification time for increasing numbers of relays was: $0.12 s$, $0.20 s$, $2.56 s$, $8.24 s$, $28.0 s$, $549 s$, \TO, \TO, \TO (compare to the numbers in the second column of Figure~\ref{fig:benchmark}). Empirical evidence suggests that our verification, while slower for small numbers of nodes, scales much better for larger numbers of nodes.

@@

To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs. The results are summarized in Figure~\ref{fig:effort} and sketched below. In the table, for each modification, we report how much of the model must be modified. We report on four aspects of the system: which transition systems making up the full model are modified (as described in Section~\ref{sec:model}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to.

@@

Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time. As well, a ``receive window'' is defined at which messages from non-faulty nodes should be received. Messages received outside the window are marked as coming from manifest-faulty senders. To implement a time-triggered model, we create three definitions: encoding nondeterministic message delay, and two definitions that are helpers in invariants. The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults. Additionally, the ASM definition is modified to track the times in the calendar, not just the messages. Two new invariants calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows. Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending manifestly-faulty messages.

@@

The Oral Messages algorithm and its variants have a long history of formal verification, making it a good test-case for comparing and contrasting verification tools~\cite{pvs}. The Oral Messages algorithm was presented in the landmark paper by Lamport, Shostak, and Pease~\cite{om}. OM(1) was subsequently verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}. Also in ACL2, an implementation of a circuit design to implement OM(1) is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail. A refinement-based verification approach is used, and OM(1) is specialized to a four-node system. It appears that a general purpose voting algorithm (like Fast MJRTY) is not verified as part of the work. A hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}. Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{Lincoln-Rushby}.

Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults, specialized for three or four nodes~\cite{om1-synth}. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. The model is efficient for up to five nodes, but results are not presented beyond that. Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of OM(1) as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The goal of the language is to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, a model-checking descendent to SAL that will be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}

>>>>

\usepackage{ifthen}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{alltt}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{multirow}
\usepackage{multicol}

@@

\newcommand{\OM}[1]{\ensuremath{\mathrm{OM}(#1)}\xspace}
\newcommand{\OMH}{\ensuremath{\mathrm{OMH}}\xspace}
\newcommand{\Id}{\ensuremath{\mathrm{Id}}\xspace}
\newcommand{\Msg}{\ensuremath{\mathrm{Msg}}\xspace}
\newcommand{\ERR}{\ensuremath{\mathrm{ERR}}\xspace}
\newcommand{\TO}{\ensuremath{\bm{\mathrm{T}}}\xspace}

@@

With this machinery, we can define the semantics of faults by constraining the relationship between a message broadcast and the values received by the recipients. For a nonfaulty process that broadcasts, every recipient receives the sent message, and for symmetric faults, there is no requirement that the messages sent are the ones received, only that every recipient receives the same value:
\noindent

\begin{center}
\begin{tabular}{l|l}
\begin{minipage}[t]{0.5\linewidth}
\[\begin{aligned}
&nonfaulty\_constraint =\\
  &\quad \forall i, j \in \Id, t \in \mathbb{R}.\\
  &\quad\quad\quad faults(i) = none\\
  &\quad\quad \textnormal{ implies } recv(i, j, t) = send(i, j, t)
\end{aligned}\]
\end{minipage}&
\begin{minipage}[t]{0.5\linewidth}
\[\begin{aligned}
&sym\_constraint =\\
  &\quad \forall i, j, k \in \Id, t \in \mathbb{R}.\\
  &\quad\quad\quad (\phantom{\textnormal{and }} faults(i) = sym\\
  &\quad\quad\quad \phantom{(}\textnormal{and } broadcast(i, \{j, k\}, t, m))\\
  &\quad\quad \textnormal{ implies } recv(i, j, t) = recv(i, k, t)
\end{aligned}\]
\end{minipage}
\end{tabular}
\end{center}

\noindent
Byzantine faults are left completely unconstrained.

@@

As a point of comparison, Rushby presents an elegant model of the simpler \OM algorithm, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}. The model is at the algorithmic level. For small values, the verification is much faster, probably due to making only one call to SAL. However, for six relays and two receivers, it takes 449~seconds and timeouts for seven relays and two receivers.

@@

To demonstrate the modularity of the model and verification approach, in this section, we explore variants to the model and report the effort required to implement the modifications and repair the proofs. The results are summarized in Figure~\ref{fig:effort} and sketched below. In the table, for each modification, we report how much of the model must be modified. We report on four aspects of the system: which transition systems making up the full model are modified (as described in Section~\ref{sec:model}), how many definitions have to be added or modified, the number of invariants that have to be added or modified, and which invariant classes (as defined in Section~\ref{sec:invariants}) those lemmas belong to. We modify the implementation along the axes of faults, time, and local node behavior.

\lee{no paragraph on no faults in table?}

@@

Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time. As well, a ``receive window'' is defined at which messages from non-faulty nodes should be received. Messages received outside the window are marked as coming from manifest-faulty senders. To implement a time-triggered model, we create three definitions: encoding nondeterministic message delay, and two definitions that are helpers in invariants. The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults. Additionally, the ASM definition is modified to track the times in the calendar, not just the messages. Two new invariants calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows. Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending benign messages.

@@

The Oral Messages algorithm and its variants and its variants have a long history of formal verification. \OM{1} was verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}. Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail. A refinement-based verification approach is used, and \OM{1} is specialized to a four-node system. It appears that a general purpose voting algorithm (like Fast MJRTY) is not verified as part of the work. A hybrid fault-model and modified Oral Messages algorithm was developed by Thambidurai and Park~\cite{hybrid}. Using the PVS interactive theorem-prover~\cite{pvs}, Lincoln and Rushby demonstrate a flaw in Thambidurai and Park's algorithm and present a fix to it~\cite{Lincoln-Rushby}.

Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults, specialized for three or four nodes~\cite{om1-synth}. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. The model is efficient for up to five nodes, but results are not presented beyond that. Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The goal of the language is to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
